---
title: "S5a_AHL flip"
author: "Prashant Kalvapalle"
date: "18 March 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal of the analysis

We want to test by qPCR if AHL is flipping any constructs. Using this file to normalize copy #s to that of the Backbone

## Running source codes

```{r cars}
setwd("../..")
source('general_functions.R')
source('analysis.R')
```


## Transforming data and normalization

1. Select only relevant data from the results
2. Data divided into 3 tibbles for each primer-pair (also called as Target)


3. Ratio of Flipped/Backbone and Unflipped/backbone are calculated

4. data is combined into 1 data frame
5. And plotted in next section

```{r}
sel <- results_abs %>% select(`Sample Name`,assay_variable,`Primer pair`,Target,`Copy #`)
# sel_grouped <- sel %>% group_by(Target) %>% mutate(id = 1:n())
# sel_spread <- sel_grouped %>% spread(Target,`Copy #`)


sel_b <- sel %>% filter(Target == 'Backbone')
sel_f <- sel %>% filter(Target == 'Flipped'); sel_u <- sel %>% filter(Target == 'Unflipped');

sel_f %<>% mutate("Normalized copy #" = sel_f$`Copy #`/sel_b$`Copy #`); 
sel_u %<>% mutate("Normalized copy #" = sel_u$`Copy #`/sel_b$`Copy #`);

results_ratio <- bind_rows(sel_f, sel_u)  
```

## Including Plots

Plots here

```{r pressure, echo=FALSE}
  
plt <- results_ratio %>% ggplot(aes(x = `assay_variable`, y = `Normalized copy #`, color = Target)) +   # plotting
  scale_y_log10(  # logscale for y axis with tick marks
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x) )
  )

plt <- plt + geom_point() +
    theme_classic() + scale_color_brewer(palette="Set1") + 
    theme(plot.title = element_text(hjust = 0.5),axis.text.x = element_text(angle = 90, hjust = 1, vjust = .3)) + 
    ggtitle(title_name) + facet_wrap(~`Sample Name`, scales = 'free_x')

print(plt)
# ggsave('S5a_AHL flip_normalized.png')

```